(Re)building cooperation: Effects of a cognitive intervention on cooperative behavior in games

### 2023-09-25 article.Rmd
revision 25sep2023

abstract: |
  Social trust is an important building block of strong social bonds, and its absence is a risk factor for social dysfunction. As such, interventions to foster and strengthen trust-based cooperation are highly desirable. Using the Repeated Trust Game paradigm, we assess the effectiveness of a cognitive intervention derived from Dialectical Behavior Therapy. The intervention's goal was to repair the potential breakdown of cooperation from a pre-programmed, one-off defection by the opponent. Over two games, participants are given the role of the trustee and face what they believe are two different players. In between games, they either receive a brief cognitive intervention or not. Post-intervention, participants showed more cooperative behavior both before and after defection by the opponent. Analysing participants' actions with a hidden Markov model shows that participants in the intervention group had a lower probability of transitioning to non cooperative states, and a higher proportion of cooperative states, than participants in the control group. This is consistent with participants inferring from the intervention that pro-social and trustworthy behavior may generally provide more beneficial outcomes to them in the long run.

bibliography: "bib/cogIntervention.bib"
biblio-style: spphys
output:
  bookdown::pdf_document2:
   toc: false
---
\small  
^1^ *Department of Experimental Psychology, Division of Psychology and Language Sciences, UCL*.  
^2^ *Division of Psychiatry and Max Planck Centre for Computational Psychiatry and Ageing Research, Queen Square Institute of Neurology, UCL.*

@@

```{r HMMPanels, fig.cap=" We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round Repeated Trust Game against human trustees. From the fitted HMM, we get the distribution of investments by the artificial investor agent conditional on its latent state as shown in Panel A. The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel B. Each plot in Panel B represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1. The policy in each state in Panel A and the way the agent transitions between states as defined in Panel B characterises the artificial agent's behavior in the Repeated Trust Game.", fig.align="center"}

@@

@@

```{r, include=FALSE}
pctRet_all <- afex::afex_plot(mod_returns_pct, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin, 
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("All rounds percentage returns") + 
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "bottom")

@@

```{r, include=FALSE, cache=TRUE}
mod_invs <- mixed( investment ~ gameNum.f*condition.f  + (1 + gameNum.f| playerId), avg_ret_df, method="KR")
summary(mod_invs)
anova(mod_invs)

@@

inv_all <- afex_plot(mod_invs, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), 
            legend_title = "Condition") + 
  theme_bw() + 
  ggtitle("All rounds investments") +
  labs(y = "HMM Investment", x = "Game") + 
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "bottom")

@@

mod_returns_pre <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| playerId), pre_int_data, REML= TRUE, method="KR")
anova(mod_returns_pre)
summary(mod_returns_pre)

pairs_pre_int <- pairs(emmeans::emmeans(mod_returns_pre, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400, reverse=TRUE))

```
```{r, include=F}
saveRDS(summary(mod_returns_pre), "data/mod_returns_pre")
```

@@

```{r, include=F}
pctRet_pre <- afex_plot(mod_returns_pre, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), 
            legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("Pre defection trials percentage returns")+
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "bottom")
```

@@

```{r, include=FALSE}
pctRet_post <- afex_plot(mod_postcoax_trust, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), 
            legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("Post defection trials percentage returns")+
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "bottom")
```

```{r violinPanels, fig.cap="Participants in the Intervention condition returned higher proportions of the multiplied investment received in the second game compared to the first game over all ronds (Panel A). This was also the case both before (Panel C) and after (Panel D) the pre-programmed defection by the HMM investor. Those in the control conditions returned similar amounts between games before the defection, and lower amounts in the second game after the defection by the investor. Investments by the HMM agent (Panel B) were higher in the second game compared to the first game across conditions. All panels show marginal means and distributions of either investments or percentage returns across participants by Game and Condition", fig.align="center", fig.width=6, fig.height = 6}

@@

```{r emoPlotCoax, fig.cap="Self-reported emotion valence and arousal as well as investment z-scores for each round of the Repeated Trust Game averaged across participants in the intervention condition only. The participants' emotion self-reports for the investor's defection were similar before and after the intervention.",fig.align="center", fig.width=7, fig.height = 4}

@@

@@

@@

```{r, include=F}
# define a response class which only contains the standard slots, no additional slots
setClass("discgaus", contains="response", slots=c(breaks="numeric"))

@@

setMethod("discgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("discgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks)
            mod
          }
)

@@

setMethod("dens","truncdiscgaus",
          function(object,log=FALSE) {
            breaks <- c(object@min, object@breaks[object@breaks > object@min & object@breaks < object@max], object@max)
            p <- pnorm(breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(breaks[-length(breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
            p <- p/sum(p)
            if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
          }
)

@@

setMethod("truncdiscgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20, ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("truncdiscgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks, min=min, max=max)
            mod
          }
)

order_mod_gauss <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation 
    sum[i] = mod@response[[i]][[1]]@parameters$mu
  }
  # reordering the states
  mod <- label_switch(mod, rank(sum))
  return(mod)
}

@@

setMethod("dens","vtdgaus",
  function(object,log=FALSE) {
      # determine cuts for pnorm based on 
      p <- pnorm(sapply(object@y + .5*(1/object@yield),function(x) min(x,1.001)), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(sapply(object@y - .5*(1/object@yield),function(x) max(x,-0.001)), mean = object@parameters$mu, sd = object@parameters$sigma)
      norm <- (pnorm(1 + .001, mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(-0.001, mean = object@parameters$mu, sd = object@parameters$sigma))
      p <- p/norm
      # alternatively, normalize factor based on size of cuts
      #p <- pnorm(object@y + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@y - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
      #p <- p/(pnorm(1 + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(0 - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma))
    # probability when yield == 0 is always 1 
      p[object@yield == 0] <- 1
      if(log) return(log(p)) else return(p)
    }
)

@@

  rModels <- rep(list(list(vtdgaus(y=avg_ret_df$ret_pct_0,yield=3*avg_ret_df$investment))),each=i)

@@

}
```

# fittedSimple <- fittedCoax  <- fittedCtrl  <- fittedFullCtrst  <- fittedPrePost  <- list()
# # increase default in multistart from 10 to 20. or set the same seed for every model
# 
# set.seed(20221028)
# for(i in n_state_min:n_state_max) {
# 
#   fittedSimple[[i]] <- multistart(simple_HMMs[[i]],nstart=20)
# }
# save(fittedSimple, file="fittedSimple.RData")
# ##############################
# 
# set.seed(20221028)
# for(i in n_state_min:n_state_max) {
#   fittedCtrl[[i]] <- multistart(ctrl_HMMs[[i]],nstart=20)
# }
# save(fittedCtrl, file="fittedCtrl.RData")
# ##############################
# 
# fittedCoax <- list()
# set.seed(20230918)
# for(i in n_state_min:n_state_max) {
#   fittedCoax[[i]] <- multistart(coax_HMMs[[i]],nstart=20)
# }
# save(fittedCoax, file="fittedCoax.RData")
# ##############################
# 
# set.seed(20221028)
# for(i in n_state_min:n_state_max) {
#   fittedPrePost[[i]] <- multistart(prePost_HMMs[[i]],nstart=20)
# }
# save(fittedPrePost, file="fittedPrePost.RData")
# #############################
# 
# 
# 
# for(i in n_state_min:n_state_max) {
#   fittedFullCtrst[[i]] <- multistart(fullCtrst_HMMs[[i]],nstart=20)
# }
# save(fittedFullCtrst, file="fittedFullCtrst.RData")
# ###############################

@@

load("data/fittedSimple.RData")
load("data/fittedCtrl.RData")
load("data/fittedCoax.RData")
load("data/fittedFullCtrst.RData")
load("data/fittedPrePost.RData")

load("data/newCtrl.RData")
load("data/newCoax.RData")
load("data/newPrePost.RData")
load("data/newFull.RData")

@@

```{r}

@@

@@

>>>>

abstract: |
  Social trust is an important building block of strong social bonds, and its absence is a risk factor for social dysfunction. As such, interventions to foster and strengthen trust-based cooperation are highly desirable. Using the Repeated Trust Game paradigm, we assess the effectiveness of a cognitive intervention derived from Dialectical Behavior Therapy. The intervention's goal was to repair the potential breakdown of cooperation from a pre-programmed, one-off defection by the opponent. Over two games, participants are given the role of the trustee and face what they believe are two different players. In between games, they either receive a brief cognitive intervention or not. Post-intervention, participants showed more cooperative behavior both before and after defection by the opponent. Analysing participants' actions with a hidden Markov model shows that participants in the intervention group had a higher proportion of cooperative states than participants in the control group. This is consistent with participants inferring from the intervention that pro-social and trustworthy behavior may generally provide more beneficial outcomes to them in the long run.

bibliography: "bib/cogIntervention.bib"
biblio-style: spphys

output:
  bookdown::pdf_document2:
   toc: false

---
\small  
^1^ *Department of Experimental Psychology, Division of Psychology and Language Sciences, UCL*.  
^2^ *Division of Psychiatry and Max Planck Centre for Computational Psychiatry and Ageing Research, Queen Square Institute of Neurology, UCL.*

@@

```{r HMMPanels, fig.cap="We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round Repeated Trust Game against human trustees. From the fitted HMM, we get the distribution of investments by the artificial investor agent conditional on its latent state as shown in Panel A. The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel B. Each plot in Panel B represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1. The policy in each state in Panel A and the way the agent transitions between states as defined in Panel B characterises the artificial agent's behavior in the Repeated Trust Game.", fig.align="center"}

@@

```{r modAllReturns, include=FALSE, cache=TRUE}

@@

```{r, include=F}
# Function to calculate y_position for geom_signif
calculate_y_position <- function(plot) {
  plot_data <- ggplot_build(plot)$data[[1]]
  max_y <- max(plot_data$ymax, na.rm = TRUE)
  y_position <- max_y + (max_y * 0.05) # Adding 5% of the maximum y-value as a margin
  return(y_position)
}
```

```{r, include=FALSE}
pctRet_all <- afex::afex_plot(mod_returns_pct, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin, 
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("All rounds percentage returns") + 
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 0.88, tip_length = 0.03, textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "*", y_position = 0.80, tip_length = 0.03, textsize = 3)  # Control Group

pctRet_all

@@

```{r modInvs, include=FALSE, cache=TRUE}
mod_invs <- mixed( investment ~ gameNum.f*condition.f  + (1 + gameNum.f| playerId), avg_ret_df, method="KR")
summary(mod_invs)
anova(mod_invs)

@@

inv_all <- afex_plot(mod_invs, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), 
            legend_title = "Condition") + 
  theme_bw() + 
  ggtitle("All rounds investments") +
  labs(y = "HMM Investment", x = "Game") + 
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 1, tip_length = -0.03,textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "***", y_position = 18, tip_length = 0.03,textsize = 3)   # Control Group

@@

mod_returns_pre <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| playerId), pre_int_data, REML= TRUE, method="KR")
anova(mod_returns_pre)
summary(mod_returns_pre)
saveRDS(mod_returns_pre, "data/mod_returns_pre")

pairs_pre_int <- pairs(emmeans::emmeans(mod_returns_pre, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400, reverse=TRUE))
saveRDS(pairs_pre_int, "data/pairs_pre_int")

# load(mod_returns_pre)
# load(pairs_pre_int)
```

@@

```{r, include=F}
pctRet_pre <- afex_plot(mod_returns_pre, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), 
            legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("Pre defection trials percentage returns")+
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") + 
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 0.95, tip_length = 0.03,textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "n.s", y_position = 0.85, tip_length = 0.03,textsize = 3)   # Control Group

@@

pctRet_post <- afex_plot(mod_postcoax_trust, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), 
            legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("Post defection trials percentage returns") +
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "n.s", y_position = 0.95, tip_length = 0.03,textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "***", y_position = 0.9, tip_length = 0.03,textsize = 3)   # Control Group

```{r violinPanels, fig.cap="Participants in the Intervention condition returned higher proportions of the multiplied investment received in the second game compared to the first game over all ronds (Panel A), whilst those in the Control condition sent back lower returns. Focusing on pre-defection trials (Panel C), those in the Control condition returned similar amounts between games, while returns were higher in the Intervention condition in the second game. Post-defection (Panel D), returns were similar for those in the Intervention condition while those in the Control condition returned lower amounts in the second game. Investments by the HMM agent (Panel B) were higher in the second game compared to the first game across conditions. All panels show marginal means and distributions of either investments or percentage returns across participants by Game and Condition", fig.align="center", fig.height=10, fig.width=10}

@@

```{r emoPlotCoax, fig.cap="Self-reported emotion valence and arousal as well as investment z-scores for each round of the Repeated Trust Game averaged across participants in the intervention condition only. The participants' emotion reaction measured during the investor's defection round were similar before (round 12) and after (round 13) the intervention.",fig.align="center", fig.width=7, fig.height = 4}

@@

```{r, include=FALSE, echo=FALSE, warning=F}

@@

# ordering states for a truncated discretised Gaussian
order_mod_truncdiscgaus <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation
    tpars <- getpars(mod@response[[i]][[1]])
    dmod <- truncdiscgaus(seq(0, 20), pstart = tpars, min=-0.5, max=20.5)
    sum[i] <- sum(seq(0,20)*dens(dmod))
  }
  cat("Expected values for states: ",sum, "\n")
  # reordering the states
  mod <- label_switch(mod, rank(sum, ties.method = "first"))
  return(mod)
}

# ordering states for a variable response truncated discretised Gaussian
order_mod_vtdgaus <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation
    tpars <- getpars(mod@response[[i]][[1]])
    dmod <- vtdgaus(seq(0, 1, length=61), pstart = tpars, yield=rep(60, 61))
    sum[i] <- sum(seq(0, 1, length=61)*dens(dmod))
  }
  # reordering the states
  mod <- label_switch(mod, rank(sum, ties.method = "first"))
  return(mod)
}

@@

# # define a response class which only contains the standard slots, no additional slots
# setClass("discgaus", contains="response", slots=c(breaks="numeric"))
# 
# # define a generic for the method defining the response class
# 
# setGeneric("discgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("discgaus"))
# 
# # define the method that creates the response class
# 
# setMethod("discgaus", 
#           signature(y="ANY"), 
#           function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), ...) {
#             y <- matrix(y,length(y))
#             x <- matrix(1)
#             parameters <- list()
#             npar <- 2
#             if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
#             if(!is.null(pstart)) {
#               if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
#               parameters$mu <- pstart[1]
#               parameters$sigma <- pstart[2]
#             }
#             mod <- new("discgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks)
#             mod
#           }
# )
# 
# setMethod("show","discgaus",
#           function(object) {
#             cat("Gaussian with discrete support\n")
#             cat("Parameters: \n")
#             cat("mu: ", object@parameters$mu, "\n")
#             cat("sigma: ", object@parameters$sigma, "\n")
#           }
# )
# 
# setMethod("dens","discgaus",
#           function(object,log=FALSE) {
#             p <- pnorm(object@breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@breaks[-length(object@breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
#             if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
#           }
# )
# 
# setMethod("setpars","discgaus",
#           function(object, values, which="pars", ...) {
#             npar <- npar(object)
#             if(length(values)!=npar) stop("length of 'values' must be",npar)
#             # determine whether parameters or fixed constraints are being set
#             nms <- names(object@parameters)
#             switch(which,
#                    "pars"= {
#                      object@parameters$mu <- values[1]
#                      object@parameters$sigma <- values[2]
#                    },
#                    "fixed" = {
#                      object@fixed <- as.logical(values)
#                    }
#             )
#             names(object@parameters) <- nms
#             return(object)
#           }
# )
# 
# setMethod("getpars","discgaus",
#           function(object,which="pars",...) {
#             switch(which,
#                    "pars" = {
#                      parameters <- numeric()
#                      parameters <- unlist(object@parameters)
#                      pars <- parameters
#                    },
#                    "fixed" = {
#                      pars <- object@fixed
#                    }
#             )
#             return(pars)
#           }
# )
# 
# setMethod("fit","discgaus",
#           function(object,w) {
#             if(missing(w)) w <- NULL
#             if(!is.null(w)) {
#               negLL <- function(pars) {
#                 object <- setpars(object, c(pars[1], exp(pars[2])))
#                 -sum(w*log(dens(object)))
#               }
#             } else {
#               negLL <- function(pars) {
#                 object <- setpars(object, c(pars[1], exp(pars[2])))
#                 -sum(log(dens(object)))
#               }
#             }
#             pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
#             object <- setpars(object,c(pars[1], exp(pars[2])))
#             object
#           }
# )
# 
# setClass("truncdiscgaus", contains="discgaus", slots=c(min="numeric", max="numeric"))
# 
# setMethod("dens","truncdiscgaus",
#           function(object,log=FALSE) {
#             breaks <- c(object@min, object@breaks[object@breaks > object@min & object@breaks < object@max], object@max)
#             p <- pnorm(breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(breaks[-length(breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
#             p <- p/sum(p)
#             if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
#           }
# )
# 
# setGeneric("truncdiscgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("truncdiscgaus"))
# 
# setMethod("truncdiscgaus", 
#           signature(y="ANY"), 
#           function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20, ...) {
#             y <- matrix(y,length(y))
#             x <- matrix(1)
#             parameters <- list()
#             npar <- 2
#             if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
#             if(!is.null(pstart)) {
#               if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
#               parameters$mu <- pstart[1]
#               parameters$sigma <- pstart[2]
#             }
#             mod <- new("truncdiscgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks, min=min, max=max)
#             mod
#           }
# )
# 
# 
# order_mod_gauss <- function(mod) {
#   ns <- nstates(mod)
#   sum <- rep(0.0,ns)
#   for(i in 1:ns) {
#     # Expected return in each state calculation 
#     sum[i] = mod@response[[i]][[1]]@parameters$mu
#   }
#   # reordering the states
#   mod <- label_switch(mod, rank(sum))
#   return(mod)
# }
# 
# # define a response class which only contains the standard slots, no additional slots
# setClass("vtdgaus", contains="response", slots=c(yield="numeric"))
# 
# # define a generic for the method defining the response class
# 
# setGeneric("vtdgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("vtdgaus"))
# 
# # define the method that creates the response class
# 
# setMethod("vtdgaus", 
#           signature(y="ANY"), 
#           function(y, yield, pstart=NULL,fixed=NULL, ...) {
#             y <- matrix(y,length(y))
#             x <- matrix(1)
#             parameters <- list()
#             npar <- 2
#             if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
#             if(!is.null(pstart)) {
#               if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
#               parameters$mu <- pstart[1]
#               parameters$sigma <- pstart[2]
#             } else {
#               parameters <- list(mu=.5,sigma=1)
#             }
#             mod <- new("vtdgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, yield=yield)
#             mod
#           }
# )
# 
# setMethod("show","vtdgaus",
#   function(object) {
#     cat("Gaussian with variable discrete support for percentage responses\n")
#     cat("Parameters: \n")
#     cat("mu: ", object@parameters$mu, "\n")
#     cat("sigma: ", object@parameters$sigma, "\n")
#   }
# )
# 
# setMethod("dens","vtdgaus",
#   function(object,log=FALSE) {
#       # determine cuts for pnorm based on 
#       p <- pnorm(sapply(object@y + .5*(1/object@yield),function(x) min(x,1.001)), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(sapply(object@y - .5*(1/object@yield),function(x) max(x,-0.001)), mean = object@parameters$mu, sd = object@parameters$sigma)
#       norm <- (pnorm(1 + .001, mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(-0.001, mean = object@parameters$mu, sd = object@parameters$sigma))
#       p <- p/norm
#       # alternatively, normalize factor based on size of cuts
#       #p <- pnorm(object@y + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@y - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
#       #p <- p/(pnorm(1 + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(0 - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma))
#     # probability when yield == 0 is always 1 
#       p[object@yield == 0] <- 1
#       if(log) return(log(p)) else return(p)
#     }
# )
# 
# setMethod("setpars","vtdgaus",
#           function(object, values, which="pars", ...) {
#             npar <- npar(object)
#             if(length(values)!=npar) stop("length of 'values' must be",npar)
#             # determine whether parameters or fixed constraints are being set
#             nms <- names(object@parameters)
#             switch(which,
#                    "pars"= {
#                      object@parameters$mu <- values[1]
#                      object@parameters$sigma <- values[2]
#                    },
#                    "fixed" = {
#                      object@fixed <- as.logical(values)
#                    }
#             )
#             names(object@parameters) <- nms
#             return(object)
#           }
# )
# 
# setMethod("getpars","vtdgaus",
#           function(object,which="pars",...) {
#             switch(which,
#                    "pars" = {
#                      parameters <- numeric()
#                      parameters <- unlist(object@parameters)
#                      pars <- parameters
#                    },
#                    "fixed" = {
#                      pars <- object@fixed
#                    }
#             )
#             return(pars)
#           }
# )
# 
# setMethod("fit","vtdgaus",
#           function(object,w) {
#             if(missing(w)) w <- NULL
#             if(!is.null(w)) {
#               negLL <- function(pars) {
#                 object <- setpars(object, c(pars[1], exp(pars[2])))
#                 -sum(w*log(dens(object)))
#               }
#             } else {
#               negLL <- function(pars) {
#                 object <- setpars(object, c(pars[1], exp(pars[2])))
#                 -sum(log(dens(object)))
#               }
#             }
#             pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
#             object <- setpars(object,c(pars[1], exp(pars[2])))
#             object
#           }
# )

```{r,include=F}
# define a response class which only contains the standard slots, no additional slots
setClass("discgaus", contains="response", slots=c(breaks="numeric"))

@@

setMethod("discgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            } else {
              parameters <- list(mu = 10, sigma = 3)
            }
            mod <- new("discgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks)
            mod
          }
)

@@

setMethod("dens","truncdiscgaus",
          function(object,log=FALSE) {
            breaks <- c(object@min, object@breaks[object@breaks > object@min & object@breaks < object@max], object@max)
            prec <- pnorm(object@max, mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@min, mean = object@parameters$mu, sd = object@parameters$sigma)
            if(prec < 1e-12) {
              # return uniform probabilities as precision is likely off
              p <- rep(1/(length(breaks)-1), length(object@y))
              if(log) return(log(p)) else return(p)
            } else {
              p <- pnorm(breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(breaks[-length(breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
              p <- p/sum(p)
              if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
            }
          }
)

@@

setMethod("truncdiscgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), min=-0.5, max=20.5, ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("truncdiscgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks, min=min, max=max)
            mod
          }
)

@@

setMethod("dens","vtdgaus",
  function(object,log=FALSE) {
      # determine cuts for pnorm based on 
      #p <- pnorm(sapply(object@y + .5*(1/object@yield), function(x) min(x,1.001)), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(sapply(object@y - .5*(1/object@yield),function(x) max(x,-0.001)), mean = object@parameters$mu, sd = object@parameters$sigma)
      #norm <- (pnorm(1 + .001, mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(-0.001, mean = object@parameters$mu, sd = object@parameters$sigma))
      #p <- p/norm

      # alternatively, normalize factor based on size of cuts

      prec <- pnorm(1 + .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(0 - .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma)
      # check for issues with precision, using highest yield
      if(prec < 1e-12) {
        p <- 1/(object@yield + 1)
      } else {
        p <- pnorm(object@y + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@y - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
        norm <- pnorm(1 + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(0 - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
        p <- p/norm
      }
      # probability when yield == 0 is always 1 
      p[object@yield == 0] <- 1
      if(log) return(log(p)) else return(p)
    }
)

@@

  # rModels <- rep(list(list(vtdgaus(y=avg_ret_df$ret_pct_0,yield=3*avg_ret_df$investment))),each=i)

  ###########################################################
  mus <- (1:i)*(1/(i+1))
  sigmas <- rep(1/(i+1), i)
  rModels <- list()
  for(s in 1:i) {
    rModels[[s]] <- list(vtdgaus(y=avg_ret_df$ret_pct_0,yield=3*avg_ret_df$investment, pstart=c(mus[s],sigmas[s])))
  }
  #########################################################

@@

@@

load("data/newCtrl.RData")
load("data/newCoax.RData")
load("data/newPrePost.RData")

@@

load("data/newFull.RData")

@@

##### BICs ###########
# Ensure df_bics is a data frame
df_bics <- as.data.frame(df_bics)

@@

```{r, include=F, fig.align="center"}
